---
title: "MAS 646 Initial Report:"
author: "Ana Raquel Chacin, Oliver Mazariegos, Shannon Land, Mariana Gomez Del Nogal"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 7
    number_sections: yes
  html_document:
    df_print: paged
linestretch: 1
always_allow_html: yes
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
# Load dataset
df = read.csv("cleaned_merged_heart_dataset.csv")
# Set categorical variables
## Sex 1 = Male, 0 = Female
df$sex = factor(
  df$sex,
  levels = c(0,1),
  labels = c("Female", "Male")
)
df$cp = factor(
  df$cp,
  levels = c(0,1,2,3),
  labels = c("Typical angina", "Atypical angina", "Non-anginal pain", "Asymptomatic")
)
df$fbs = factor(
  df$fbs,
  levels = c(0,1),
  labels = c("False", "True")
)
df$restecg = factor(
  df$restecg,
  levels = c(0,1,2),
  labels = c("Normal", "ST-T wave abnormality", "Left ventricular hypertrophy")
)
df$exang = factor(
  df$exang,
  levels = c(0,1),
  labels = c("No", "Yes")
)
df$slope = factor(
  df$slope,
  levels = c(0,1,2),
  labels = c("Upsloping", "Flat", "Downsloping")
)
df$ca = factor(
  df$ca,
  levels = c(0,1,2,3),
  labels = c("0", "1", "2", "3")
)
df$thal = factor(
  df$thal,
  levels = c(1,2,3),
  labels = c("Normal", "Fixed defect", "Reversible defect")
)
df$target = factor(
  df$target,
  levels = c(0,1),
  labels = c("No heart attack", "Heart attack")
)

```

# Introduction

# Dataset Description

## Categorical Variables

## Numerical Variables

# Exploratory Data Analysis

## Structure of Dataset

```{r}
str(df)
```

## Checking for Null Values

```{r}
sapply(df, function(x) sum(is.na(x)))
```


## Statistical Summary

```{r}
summary(df)
```

## Frequency Distributions

```{r, fig.height=6}
names = names(df)
par(mfrow=c(3,2))
for (i in c(1,4,5,8,10)) {
  hist(df[, i], main = names[i], xlab=names[i], col = "lightblue")
}
```

## Check for Normality

```{r, fig.height=6}
par(mfrow=c(3,2))
for (i in c(1,4,5,8,10)) {
  qqnorm(df[, i], main = names[i])
  qqline(df[, i])
}
```

## Pairwise Relationship: Numerical Predictors

```{r, fig.height=7}
par(mfrow=c(3,2))
for (i in c(1,4,5,8,10)) {
  boxplot(df[,i]~df$target,main=names[i],xlab="Target",ylab=names[i])
}
```


## Pairwise Relationship: Categorical Predictors

```{r, , fig.height=7}
colors <- c("lightblue","lightpink")
##Create percentage contingency table
par(mfrow=c(3,3))
for(i in c(2,3,6,7,9,11,12,13)){
  tab <- table(df$target,df[,i])
  prob.tab <- prop.table(tab,margin = 2)
  barplot(prob.tab,beside=TRUE,xlab=names[i],col=colors)
  if(i==2) legend(x=.7,y=0.7,legend=c("No heart attack","Heart attack"),fill=colors,bty='n')
}
```


# Model Building

## Simple Logistic Regression Model

```{r}
glmod <-glm(target~.,family="binomial",data=df)
summary(glmod)
```


## Model Evaluation

### Confusion Matrix

```{r}
## Find estimated probablities for all training data
prob <- predict(glmod,newdata=df,type="response")
pred.target <- as.factor(ifelse(prob>0.5,"Heart attack", "No heart attack"))
#create the confusion matrix
library(caret)
c.matrix <- confusionMatrix(pred.target,df$target)
c.matrix$table
```


### ROC & AUC

```{r}
library(pROC)
###Create the ROC curve
roc_glm <- roc(df$target~prob)
plot(roc_glm)
auc(roc_glm)
```


